---
title: "JPI Data Processing"
author: "CHAIN Data Team"
date: "2025-04-17"
output:
  html_document: default
  pdf_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, include = TRUE)

```

### Load necesary libraries
```{r}
library(tidyverse)
library(here)
library(lubridate)
library(kableExtra)
library(janitor)
library(table1)
library(readr)
library(ggplot2)
library(DT)
library(data.table)
library(haven)
library(scales)
```

```{r}
file_path = "/srv/shiny-server/jpi/processed_data/"
```

```{r}
#load datasets
hospital_df <- readRDS(here("data","hospital_df.rds"))
community_wet_df <- readRDS(here("data","community_wet_df.rds"))
community_stool_df <- readRDS(here("data","community_stool_df.rds"))
community_care_seeking_df <- readRDS(here("data","community_care_seeking_df.rds"))
exit_df <- readRDS(here("data","exit_df.rds"))
pps_ward_df <- readRDS(here("data","pps_ward_df.rds"))
pps_patient_df <- readRDS(here("data","pps_patient_df.rds"))
rectal_swabs_df <- readRDS(here("data","samples","rectal_swabs_df.rds"))
stool_kidms_df <- readRDS(here("data","samples","stool_kidms_df.rds"))

```
### Save core datasets
```{r}
save(community_wet_df,file = paste0(file_path, "community_wet_df.RData") )
save(community_stool_df,file = paste0(file_path, "community_stool_df.RData") )
save(community_care_seeking_df,file = paste0(file_path, "community_care_seeking_df.RData") )
save(exit_df,file = paste0(file_path, "exit_df.RData") )
save(hospital_df,file = paste0(file_path, "hospital_df.RData") )
save(stool_kidms_df,file = paste0(file_path, "stool_kidms_df.RData"))
```
##clusters
```{r}
#roka_dispensary_register <- read_csv(here("data", "roka_dispensary_register.csv"))
#roka_ocean_register <- read_csv(here("data", "roka_ocean_register.csv"))

#clusters_df <- rbind(roka_dispensary_register, roka_ocean_register)

# Columns in df1 but not in df2
#cols_only_in_df1 <- setdiff(names(roka_dispensary_register), names(roka_ocean_register))

# Columns in df2 but not in df1
#cols_only_in_df2 <- setdiff(names(roka_ocean_register), names(roka_dispensary_register))

# Display mismatched columns
#list(
  #only_in_df1 = cols_only_in_df1,
 # only_in_df2 = cols_only_in_df2
#)

```

### Data Cleaning
```{r}
hospital_df$ageyrs_adm = as.numeric(hospital_df$ageyrs_adm)

### Clean dates in "POSIXct" "POSIXt" 
community_wet_df %>% 
select(starts_with("date_")) %>% 
names()

date_vars <- c(
  "date_survey_hh","date_crf_completed_hh","date_crf_reviewed_hh","date_visit_esc","date_consent_esc","date_collect_ws","date_collect_ls", "date_collect_ls_2" ,"date_collect_cf","date_collect_ss","date_water_recieve","date_latrine_recieve","date_outdoor_recieve", "date_chickfaeces_sample_recieve","date_soil_recieve","date_soil_recieve","date_crf_reviewed_esc", "date_crf_completed_esc"  )
               

# Apply transformation to each variable in date_vars
walk(date_vars, function(var) {
  community_wet_df[[var]] <<- as.Date(format(
    as.POSIXct(as.character(zap_labels(community_wet_df[[var]])), tz = "Africa/Nairobi"),
    tz = "Africa/Nairobi" ))})

```

### Summary value boxes
```{r}
community <- as.numeric(length(which(!is.na(community_wet_df$record_id))))
hospital <- as.numeric(length(which(!is.na(hospital_df$record_id))))
exit <- as.numeric(length(which(!is.na(exit_df$record_id))))

pps_ward <- as.numeric(length(which(!is.na(pps_ward_df$record_id))))

pps_patient <- as.numeric(length(which(!is.na(pps_patient_df$record_id))))

main_value_boxes <- data.frame(community = community, hospital = hospital, exit = exit, pps_ward = pps_ward, pps_patient = pps_patient)

save(main_value_boxes,file = paste0(file_path, "main_value_boxes.RData") )

```
### 1. Hospital Reports 

### hospital enrolment curve
```{r}
# use hospital df
```

### hospital participant characteristics
```{r}
part_summary <- hospital_df %>% group_by(season_adm) %>% summarize(
  enrolled = n(),
  Male = length(which(sex_adm == "Male")),
  Male = paste0(Male, " (", as.character(round(Male/enrolled * 100, 0)), "%)"),
  Mean_Age_yrs = round(mean(ageyrs_adm, na.rm = TRUE), 1)
)

save(part_summary,file = paste0(file_path, "part_summary.RData") )

```
### Demographic characteristics 
```{r}
 ## use table 1 
## Demographic variables include age_yrs,sex_adm,axilla_temp_adm,weight_adm,length_adm,prev_adm_hosp_adm,malaria_rdt_adm,hiv_known_positive

```
### Monthly hospital enrolment 
```{r}
monthly_enrolments <- hospital_df %>%
  filter(!is.na(date_adm)) %>%
  mutate(
    Year_Month = paste0(
      year(date_adm), "-",
      lubridate::month(date_adm, label = TRUE, abbr = TRUE)
    )
  ) %>%
  group_by(Year_Month) %>%
  summarise(Enrolments = n(), .groups = "drop") %>%
  arrange(parse_date_time(Year_Month, orders = "Y-b")) %>%  # optional: sorts chronologically
  adorn_totals(where = "row")

save(monthly_enrolments,file = paste0(file_path, "monthly_enrolments.RData") )

```
### 2. Community wet Reports 
```{r}
#  study start date
study_start_date <- as.Date("2025-04-01")

households_weekly_enrol <- community_wet_df %>%
  mutate(
    # Calculate week number based on survey date
    Week = as.numeric(floor(difftime(date_survey_hh, study_start_date, units = "weeks"))) + 1,
    Week_start_date = study_start_date + ((Week - 1) * 7),  # Week start date
    Week_end_date = Week_start_date + 6,  # Week end date
    Year_Month = format(date_survey_hh, "%Y-%m")  # Year-Month
  ) %>%
  group_by(Week, Week_start_date, Week_end_date) %>%
  summarise(Households_Enrolled = n_distinct(record_id), .groups = "drop") %>%
  arrange(Week) %>%
  adorn_totals(where = "row", name = "Total", fill = "-")  # Add row totals

kbl(households_weekly_enrol,caption = "Households summary",booktabs = TRUE,
    linesep = "\\addlinespace",align = "lcccc") %>%
  kable_styling(latex_options = c("striped", "scale_down", "hold_position"),font_size = 8,
    full_width = TRUE, position = "left") %>% row_spec(0, bold = TRUE) 

save(households_weekly_enrol,file = paste0(file_path, "households_weekly_enrol.RData") )
```
### Housing charactristics 
```{r}
##  variables "main_drink_water_srce_hh",             "distance_srce_hh","toilet_clean_hh",animals_kept_hh","animal_excrement_hh"
```
### Community Queries 
```{r}
mismatched_hm <- community_wet_df %>%
  mutate(
    extracted_hm_number = str_extract(record_id, "[^-]+$") %>% as.numeric(),
    check_match = extracted_hm_number == hm_number_hh
  ) %>%
  filter(!check_match) %>%  # only show mismatches
  select(record_id, date_survey_hh, hm_number_hh, crf_completed_by_hh)

kbl(mismatched_hm,caption = "Mismatched Hms",booktabs = TRUE,
    linesep = "\\addlinespace",align = "lcccc") %>%
  kable_styling(latex_options = c("striped", "scale_down", "hold_position"),font_size = 8,
    full_width = TRUE, position = "left") %>% row_spec(0, bold = TRUE) 

save(mismatched_hm,file = paste0(file_path, "mismatched_hm.RData") )

```


### Stool sample completeness

```{r}
# Overall completeness
stool_completeness_summary <- community_stool_df %>%
  summarise(
    Target_samples = 46 * 6,
    Total_participants = n(),
    Stool_collected = sum(!is.na(date_collect_stool)),
    Not_collected = sum(is.na(date_collect_stool)),
    Percent_complete = paste0(round((Stool_collected / Total_participants) * 100, 1), "%")
  )

kbl(stool_completeness_summary,caption = "Stool summary",booktabs = TRUE,
    linesep = "\\addlinespace",align = "lcccc") %>%
  kable_styling(latex_options = c("striped", "scale_down", "hold_position"),font_size = 8,
    full_width = TRUE, position = "left") %>% row_spec(0, bold = TRUE)

save(stool_completeness_summary,file = paste0(file_path, "stool_completeness_summary.RData") )

```
### Reasons stool not collected 
```{r}
rsns_stool_not_collected <- community_stool_df %>%
  filter(sample_collected_stool == "No") %>%
  group_by(rson_no_sample_stool) %>%
  summarise(Total = n(), .groups = "drop") %>%
  mutate(
    Percent = round((Total / sum(Total)) * 100, 1)
  ) %>%
  mutate(
    Percent = paste0(Percent, "%")
  ) %>%
  adorn_totals("row", name = "Total", fill = "")

save(rsns_stool_not_collected,file = paste0(file_path, "rsns_stool_not_collected.RData"))

```
### Stool kidms queries
```{r}
stool_wrong_serial_study_id <- stool_kidms_df %>% 
  dplyr::filter(serial_study_id %in% c("1-61B-14-3W-F","	
1-120-409-W-F1","1-120-409-W-F3","1-120-409-W-F4","1-12A-153-W-F","1-12A-153-W-F2","1-12A-153-W-F3","1-12A-153-W-F5","1-12A-153-W-F6","1-12A-44-2-F","1-12G-553-2W-F","1-12G-553-2W-F","112A110W-F2","112A110W-F3","112H185W-F1",	
"112H185W-F2","112D282W-F1","	
112D282W-F2","112N175W-F2","112A287-F1337","	
112A287-F1","112A287-F2","112A287-F3","1-116G-478-W-F","1-113-428-6-W-F")) %>% 
  select(serial_study_id,pk_specimen_id,date_collect,type_name)
  
kbl(stool_wrong_serial_study_id,caption = "Wrong format of serial study id",booktabs = TRUE,
    linesep = "\\addlinespace",align = "lcccc") %>%
  kable_styling(latex_options = c("striped", "scale_down", "hold_position"),font_size = 8,
    full_width = TRUE, position = "left") %>% row_spec(0, bold = TRUE) 

write_csv(stool_wrong_serial_study_id, file = here("queries","stool_wrong_serial_study_id.csv"))

save(stool_wrong_serial_study_id,file = paste0(file_path, "stool_wrong_serial_study_id.RData"))

## Missing time_point
miss_stool_time_point<-stool_kidms_df %>% 
  dplyr::filter(!time_point == "Wet") %>% 
  select(serial_study_id,pk_specimen_id,date_collect,time_point)

kbl(miss_stool_time_point,caption = "Missing stool time point ",booktabs = TRUE,
    linesep = "\\addlinespace",align = "lcccc") %>%
  kable_styling(latex_options = c("striped", "scale_down", "hold_position"),font_size = 8,
    full_width = TRUE, position = "left") %>% row_spec(0, bold = TRUE) 

save(miss_stool_time_point,file = paste0(file_path, "miss_stool_time_point.RData"))
```

### Rectal swabs - Admission samples
```{r}
rectal_swab_redcap <- hospital_df %>% 
  select(record_id,sample_num_rectswab1_adm,sample_num_rectswab2_adm,date_rectswab_adm,rectswab_by_adm)

admission_samples_summary <- hospital_df %>%
  summarise(
    Rectal_Swab = sum(!is.na(date_rectswab_adm))
    ) %>%
  pivot_longer(
    cols = everything(),
    names_to = "Sample_Type",
    values_to = "Total_Collected"
  )
save(admission_samples_summary,file = paste0(file_path, "admission_samples_summary.RData") )

kbl(admission_samples_summary,caption = "Admission samples summary",booktabs = TRUE,
    linesep = "\\addlinespace",align = "lcccc") %>%
  kable_styling(latex_options = c("striped", "scale_down", "hold_position"),font_size = 8,
    full_width = TRUE, position = "left") %>% row_spec(0, bold = TRUE)
```
### Rectal swab queries _kidms

```{r}
rectswab1_miss <- rectal_swabs_df %>% 
  dplyr::filter(serial_study_id %in% c("1P002-W-R2","1P002-W-R2","1P003-W-R2","1P003-W-R2")) %>% 
  select(serial_study_id,type_name,date_collect,pk_specimen_id)
 
kbl(rectswab1_miss,caption = "Record ids missing rectal swab 1",booktabs = TRUE,
    linesep = "\\addlinespace",align = "lcccc") %>%
  kable_styling(latex_options = c("striped", "scale_down", "hold_position"),font_size = 8,
    full_width = TRUE, position = "left") %>% row_spec(0, bold = TRUE) 

save(rectswab1_miss,file = paste0(file_path, "rectswab1_miss.RData")) 

### wrong serial number format 
wrong_serial_study_id_format <- rectal_swabs_df %>% 
  dplyr::filter(serial_study_id %in% c("1P-023-W-R2","1P026","IP027-W-R1","IP027-W-R2
","1PO42-W-R2","1P060")) %>% 
  select(serial_study_id,type_name,date_collect,pk_specimen_id)
  
kbl(wrong_serial_study_id_format,caption = "Wrong format of serial study id",booktabs = TRUE,
    linesep = "\\addlinespace",align = "lcccc") %>%
  kable_styling(latex_options = c("striped", "scale_down", "hold_position"),font_size = 8,
    full_width = TRUE, position = "left") %>% row_spec(0, bold = TRUE) 

write_csv(wrong_serial_study_id_format, file = here("queries","wrong_serial_study_id_format.csv"))

save(wrong_serial_study_id_format,file = paste0(file_path, "wrong_serial_study_id_format.RData")) 

### Rectal swab specimen source 
rectswab_specimen_source <- rectal_swabs_df %>% 
  dplyr::filter(specimen_source %in% c("clinic","field")) %>% 
  select(serial_study_id,specimen_source,type_name,date_collect)

kbl(rectswab_specimen_source,caption = "Wrong rectal_swab specimen source",booktabs = TRUE,
    linesep = "\\addlinespace",align = "lcccc") %>%
  kable_styling(latex_options = c("striped", "scale_down", "hold_position"),font_size = 8,
    full_width = TRUE, position = "left") %>% row_spec(0, bold = TRUE) 

write_csv(rectswab_specimen_source, file = here("queries","rectswab_specimen_source.csv"))

save(rectswab_specimen_source,file = paste0(file_path, "rectswab_specimen_source.RData")) 

## Rectal swabs missing date recieved 

rectswab_date_recieved_miss <- rectal_swabs_df %>%
  dplyr::filter(is.na(date_received)) %>% 
  select(serial_study_id, type_name,pk_specimen_id,date_collect, date_received)
  
kbl(rectswab_date_recieved_miss,caption = "Rectswab missing date_recieved in the lab",booktabs = TRUE,
    linesep = "\\addlinespace",align = "lcccc") %>%
  kable_styling(latex_options = c("striped", "scale_down", "hold_position"),font_size = 8,
    full_width = TRUE, position = "left") %>% row_spec(0, bold = TRUE) 

write_csv(rectswab_date_recieved_miss, file = here("queries","rectswab_date_recieved_miss.csv"))

save(rectswab_date_recieved_miss,file = paste0(file_path, "rectswab_date_recieved_miss.RData"))

## Rectal swabs missing time recieved 

rectswab_time_recieved_miss <- rectal_swabs_df %>%
  dplyr::filter(is.na(time_received)) %>% 
  select(serial_study_id, type_name,pk_specimen_id,date_collect, time_received)
  
kbl(rectswab_time_recieved_miss,caption = "Rectswab missing time_recieved in the lab",booktabs = TRUE,
    linesep = "\\addlinespace",align = "lcccc") %>%
  kable_styling(latex_options = c("striped", "scale_down", "hold_position"),font_size = 8,
    full_width = TRUE, position = "left") %>% row_spec(0, bold = TRUE) 

write_csv(rectswab_time_recieved_miss, file = here("queries","rectswab_time_recieved_miss.csv"))

save(rectswab_time_recieved_miss,file = paste0(file_path, "rectswab_time_recieved_miss.RData"))

```
### Environmental samples  

```{r}
# environmental_sample_summary <- community_wet_df %>%
#   summarise(
#     Total_homesteads = sum(!is.na(date_survey_hh), na.rm = TRUE), 
#     Water_samples = sum(!is.na(date_collect_ws), na.rm = TRUE),
#     Soil_samples = sum(!is.na(date_collect_ss), na.rm = TRUE),
#     Latrine_samples = sum(!is.na(date_collect_ls), na.rm = TRUE),
#     Outdoor_samples = sum(!is.na(date_collect_ls_2), na.rm = TRUE),
#     Chicken_faeces_samples = sum(!is.na(date_collect_cf), na.rm = TRUE),
#     
#     Total_samples = Water_samples + Soil_samples + Latrine_samples + Outdoor_samples + Chicken_faeces_samples
#   )
# save(environmental_sample_summary,file = paste0(file_path, "environmental_sample_summary.RData") )


```
### Environmental samples 2

```{r}
environmental_sample_summary_2 <- community_wet_df %>%
  summarise(
    Total_homesteads = sum(!is.na(date_survey_hh), na.rm = TRUE), 
    
    Water_samples = paste0(
      sum(!is.na(date_collect_ws), na.rm = TRUE), " (",
      round(sum(!is.na(date_collect_ws), na.rm = TRUE) / Total_homesteads * 100, 1), "%)"
    ),
    
    Soil_samples = paste0(
      sum(!is.na(date_collect_ss), na.rm = TRUE), " (",
      round(sum(!is.na(date_collect_ss), na.rm = TRUE) / Total_homesteads * 100, 1), "%)"
    ),
    
    Latrine_samples = paste0(
      sum(!is.na(date_collect_ls), na.rm = TRUE), " (",
      round(sum(!is.na(date_collect_ls), na.rm = TRUE) / Total_homesteads * 100, 1), "%)"
    ),
    
    Outdoor_samples = paste0(
      sum(!is.na(date_collect_ls_2), na.rm = TRUE), " (",
      round(sum(!is.na(date_collect_ls_2), na.rm = TRUE) / Total_homesteads * 100, 1), "%)"
    ),
    
    Chicken_faeces_samples = paste0(
      sum(!is.na(date_collect_cf), na.rm = TRUE), " (",
      round(sum(!is.na(date_collect_cf), na.rm = TRUE) / Total_homesteads * 100, 1), "%)"
    )
  ) %>%
  pivot_longer(
    cols = -Total_homesteads,
    names_to = "Sample_Type",
    values_to = "Sample_Count_%"
  )
save(environmental_sample_summary_2,file = paste0(file_path, "environmental_sample_summary_2.RData") )
```
### Reasons Latrine sample not collected 
```{r}
rsns_latrine_not_collected <- community_wet_df %>%
  filter(latrine_sample_collected_ls == "No") %>%
  group_by(rson_no_sample_ls) %>%
  summarise(Total = n(), .groups = "drop") %>%
  mutate(
    Percent = round((Total / sum(Total)) * 100, 1)
  ) %>%
  mutate(
    Percent = paste0(Percent, "%")
  )

save(rsns_latrine_not_collected,file = paste0(file_path, "rsns_latrine_not_collected.RData"))
```

```{r}
rsns_latrine_not_collected <- community_wet_df %>%
  filter(latrine_sample_collected_ls == "No") %>% 
  select(record_id,latrine_sample_collected_ls,rson_no_sample_ls,outdoor_sample_collected_ls,rson_no_sample_od_ls,latrine_sample_by_ls,outdoor_sample_by_ls)

write_csv(rsns_latrine_not_collected, here("data","rsns_latrine_not_collected.csv"))
```
### PPS Ward

```{r}
# First report: Summary of wards
pps_ward_summary <- pps_ward_df %>%
  filter(!is.na(ward_paeds_pps)) %>%
  group_by(ward_paeds_pps) %>%
  summarise(
    total_enrolled_ward = n(),
    .groups = "drop"
  ) %>%
  rename(Ward_Name = ward_paeds_pps)

# Second report: Summary of patients
pps_patient_summary <- pps_patient_df %>%
  filter(!is.na(ward_name_pps)) %>%
  group_by(ward_name_pps) %>%
  summarise(
    Total_patients = n(),
    .groups = "drop"
  ) %>%
  rename(Ward_Name = ward_name_pps)

save(pps_patient_summary,file = paste0(file_path, "pps_patient_summary.RData"))
```

```{r}




```

