---
title: "JPI Data Processing"
author: "CHAIN Data Team"
date: "2025-04-17"
output:
  html_document: default
  pdf_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, include = TRUE)

```

### Load necesary libraries
```{r}
library(tidyverse)
library(here)
library(lubridate)
library(kableExtra)
library(janitor)
library(table1)
library(readr)
library(ggplot2)
library(DT)
library(data.table)
library(haven)
library(scales)
```

```{r}
file_path = "/srv/shiny-server/jpi/processed_data/"
```

```{r}
#load datasets
hospital_df <- readRDS(here("data","hospital_df.rds"))
hospital_paeds <- readRDS(here("data","hospital_paeds.rds"))
hospital_adults <- readRDS(here("data","hospital_adults.rds"))
community_wet_df <- readRDS(here("data","community_wet_df.rds"))
community_stool_df <- readRDS(here("data","community_stool_df.rds"))
community_care_seeking_df <- readRDS(here("data","community_care_seeking_df.rds"))
exit_df <- readRDS(here("data","exit_df.rds"))
pps_ward_df <- readRDS(here("data","pps_ward_df.rds"))
pps_patient_df <- readRDS(here("data","pps_patient_df.rds"))

## KIDMS 
rectal_swabs_kidms_df <- readRDS(here("data","kidms_samples","rectal_swabs_kidms_df.rds"))

community_stool_kidms_df <- readRDS(here("data","kidms_samples","community_stool_kidms_df.rds"))

```
### Data Cleaning
```{r}
### Clean dates in "POSIXct" "POSIXt"
hospital_df$ageyrs_adm = as.numeric(hospital_df$ageyrs_adm)
hospital_df <- hospital_df %>% 
  mutate(
    cohort = if_else(ageyrs_adm <= 5, "Paed", 
                     if_else(ageyrs_adm >= 18, "Adult", "")),
    )
save(hospital_df,file = paste0(file_path, "hospital_df.RData"))
```

```{r}
### Clean dates in "POSIXct" "POSIXt" 
community_wet_df %>% 
select(starts_with("date_")) %>% 
names()

date_vars <- c(
  "date_survey_hh","date_crf_completed_hh","date_crf_reviewed_hh","date_visit_esc","date_consent_esc","date_collect_ws","date_collect_ls", "date_collect_ls_2" ,"date_collect_cf","date_collect_ss","date_water_recieve","date_latrine_recieve","date_outdoor_recieve", "date_chickfaeces_sample_recieve","date_soil_recieve","date_soil_recieve","date_crf_reviewed_esc", "date_crf_completed_esc"  )
              
# Apply transformation to each variable in date_vars
walk(date_vars, function(var) {
  community_wet_df[[var]] <<- as.Date(format(
    as.POSIXct(as.character(zap_labels(community_wet_df[[var]])), tz = "Africa/Nairobi"),
    tz = "Africa/Nairobi" ))})
save(community_wet_df,file = paste0(file_path, "community_wet_df.RData") )
```

```{r}
### Clean dates in "POSIXct" "POSIXt" 
community_stool_df %>% 
select(starts_with("date_")) %>% 
names()

date_vars <- c(
  "date_stool","date_consent_stool","date_collect_stool","date_stool_passing" ,"date_crf_completed_stool","date_crf_reviewed_stool")
               

# Apply transformation to each variable in date_vars
walk(date_vars, function(var) {
  community_stool_df[[var]] <<- as.Date(format(
    as.POSIXct(as.character(zap_labels(community_stool_df[[var]])), tz = "Africa/Nairobi"),
    tz = "Africa/Nairobi" ))})
save(community_stool_df,file = paste0(file_path, "community_stool_df.RData") )
```

### Save core datasets
```{r}
save(community_care_seeking_df,file = paste0(file_path, "community_care_seeking_df.RData") )
save(exit_df,file = paste0(file_path, "exit_df.RData") )

save(rectal_swabs_kidms_df,file = paste0(file_path, "rectal_swabs_kidms_df.RData") )

save(community_stool_kidms_df,file = paste0(file_path, "community_stool_kidms_df.RData"))
```
##clusters
```{r}
#roka_dispensary_register <- read_csv(here("data", "roka_dispensary_register.csv"))
#roka_ocean_register <- read_csv(here("data", "roka_ocean_register.csv"))

#clusters_df <- rbind(roka_dispensary_register, roka_ocean_register)

# Columns in df1 but not in df2
#cols_only_in_df1 <- setdiff(names(roka_dispensary_register), names(roka_ocean_register))

# Columns in df2 but not in df1
#cols_only_in_df2 <- setdiff(names(roka_ocean_register), names(roka_dispensary_register))

# Display mismatched columns
#list(
  #only_in_df1 = cols_only_in_df1,
 # only_in_df2 = cols_only_in_df2
#)

```

### Summary value boxes
```{r}
community <- as.numeric(length(which(!is.na(community_wet_df$record_id))))
hospital <- as.numeric(length(which(!is.na(hospital_df$record_id))))
paeds <- as.numeric(length(which(!is.na(hospital_paeds$record_id))))
adults <- as.numeric(length(which(!is.na(hospital_adults$record_id))))
exit <- as.numeric(length(which(!is.na(exit_df$record_id))))
pps_ward <- as.numeric(length(which(!is.na(pps_ward_df$record_id))))
pps_patient <- as.numeric(length(which(!is.na(pps_patient_df$record_id))))

main_value_boxes <- data.frame(community = community, hospital = hospital,paeds = paeds, adults = adults, exit = exit, pps_ward = pps_ward, pps_patient = pps_patient)

save(main_value_boxes,file = paste0(file_path, "main_value_boxes.RData") )

```
### 1. Hospital Reports 

### hospital enrolment curve
```{r}
# use hospital df
```

### hospital participant characteristics
```{r}
part_summary <- hospital_df %>% group_by(season_adm) %>% summarize(
  enrolled = n(),
  Male = length(which(sex_adm == "Male")),
  Male = paste0(Male, " (", as.character(round(Male/enrolled * 100, 0)), "%)"),
  Mean_Age_yrs = round(mean(ageyrs_adm, na.rm = TRUE), 1)
)

save(part_summary,file = paste0(file_path, "part_summary.RData") )

```
### Demographic characteristics 
```{r}
 ## use table 1 
## Demographic variables include age_yrs,sex_adm,axilla_temp_adm,weight_adm,length_adm,prev_adm_hosp_adm,malaria_rdt_adm,hiv_known_positive

```
### Monthly hospital enrolment 
```{r}
monthly_enrolments <- hospital_df %>%
  filter(!is.na(date_adm)) %>%
  mutate(
    Year_Month = paste0(
      year(date_adm), "-",
      lubridate::month(date_adm, label = TRUE, abbr = TRUE)
    )
  ) %>%
  group_by(Year_Month) %>%
  summarise(Enrolments = n(), .groups = "drop") %>%
  arrange(parse_date_time(Year_Month, orders = "Y-b")) %>%  # optional: sorts chronologically
  adorn_totals(where = "row")

save(monthly_enrolments,file = paste0(file_path, "monthly_enrolments.RData") )

```
### 2. Community wet Reports 
```{r}
#  study start date
study_start_date <- as.Date("2025-04-01")

households_weekly_enrol <- community_wet_df %>%
  mutate(
    # Calculate week number based on survey date
    Week = as.numeric(floor(difftime(date_survey_hh, study_start_date, units = "weeks"))) + 1,
    Week_start_date = study_start_date + ((Week - 1) * 7),  # Week start date
    Week_end_date = Week_start_date + 6,  # Week end date
    Year_Month = format(date_survey_hh, "%Y-%m")  # Year-Month
  ) %>%
  group_by(Week, Week_start_date, Week_end_date) %>%
  summarise(Households_Enrolled = n_distinct(record_id), .groups = "drop") %>%
  arrange(Week) %>%
  adorn_totals(where = "row", name = "Total", fill = "-")  # Add row totals

kbl(households_weekly_enrol,caption = "Households summary",booktabs = TRUE,
    linesep = "\\addlinespace",align = "lcccc") %>%
  kable_styling(latex_options = c("striped", "scale_down", "hold_position"),font_size = 8,
    full_width = TRUE, position = "left") %>% row_spec(0, bold = TRUE) 

save(households_weekly_enrol,file = paste0(file_path, "households_weekly_enrol.RData") )
```
### Housing charactristics 
```{r}
##  variables "main_drink_water_srce_hh",             "distance_srce_hh","toilet_clean_hh",animals_kept_hh","animal_excrement_hh"
```
### Community Queries 
```{r}
mismatched_hm <- community_wet_df %>%
  mutate(
    extracted_hm_number = str_extract(record_id, "[^-]+$") %>% as.numeric(),
    check_match = extracted_hm_number == hm_number_hh
  ) %>%
  filter(!check_match) %>%  # only show mismatches
  select(record_id, date_survey_hh, hm_number_hh, crf_completed_by_hh)

kbl(mismatched_hm,caption = "Mismatched Hms",booktabs = TRUE,
    linesep = "\\addlinespace",align = "lcccc") %>%
  kable_styling(latex_options = c("striped", "scale_down", "hold_position"),font_size = 8,
    full_width = TRUE, position = "left") %>% row_spec(0, bold = TRUE) 

save(mismatched_hm,file = paste0(file_path, "mismatched_hm.RData") )

```

### Community Stool sample completeness

```{r}
# Overall completeness
stool_completeness_summary <- community_stool_df %>%
  summarise(
    Target_samples = 120 * 6,
    Total_participants = n(),
    Stool_collected = sum(!is.na(date_collect_stool)),
    Not_collected = sum(is.na(date_collect_stool)),
    Percent_complete = paste0(round((Stool_collected / Total_participants) * 100, 1), "%")
  )

kbl(stool_completeness_summary,caption = "Stool summary",booktabs = TRUE,
    linesep = "\\addlinespace",align = "lcccc") %>%
  kable_styling(latex_options = c("striped", "scale_down", "hold_position"),font_size = 8,
    full_width = TRUE, position = "left") %>% row_spec(0, bold = TRUE)

save(stool_completeness_summary,file = paste0(file_path, "stool_completeness_summary.RData") )

```

### Reasons stool not collected 
```{r}
rsns_stool_not_collected <- community_stool_df %>%
  filter(sample_collected_stool == "No") %>%
  group_by(rson_no_sample_stool) %>%
  summarise(Total = n(), .groups = "drop") %>%
  mutate(
    Percent = round((Total / sum(Total)) * 100, 1)
  ) %>%
  mutate(
    Percent = paste0(Percent, "%")
  ) %>%
  adorn_totals("row", name = "Total", fill = "")

save(rsns_stool_not_collected,file = paste0(file_path, "rsns_stool_not_collected.RData"))

```
### Stool Redcap - Queries 
```{r}
missing_stool_information <- community_stool_df %>% 
  dplyr::filter(is.na(date_stool))
write_csv(missing_stool_information, file = here("queries","missing_stool_information.csv"))

save(missing_stool_information,file = paste0(file_path, "missing_stool_information.RData"))
```
#### Early dates - Before start date 2025-04-01
```{r}

start_date <- as.Date("2025-03-31")

early_dates_combined <- bind_rows(
  community_stool_df %>%
    filter(!is.na(date_stool) & as.Date(date_stool) < start_date) %>%
    select(record_id, date_value = date_stool,redcap_repeat_instance,crf_completed_by_stool) %>%
    mutate(date_column = "date_stool"),

  community_stool_df %>%
    filter(!is.na(date_consent_stool) & as.Date(date_consent_stool) < start_date) %>%
    select(record_id, date_value = date_consent_stool,redcap_repeat_instance,crf_completed_by_stool) %>%
    mutate(date_column = "date_consent_stool"),

  community_stool_df %>%
    filter(!is.na(date_collect_stool) & as.Date(date_collect_stool) < start_date) %>%
    select(record_id, date_value = date_collect_stool,redcap_repeat_instance,crf_completed_by_stool) %>%
    mutate(date_column = "date_collect_stool"),

  community_stool_df %>%
    filter(!is.na(date_stool_passing) & as.Date(date_stool_passing) < start_date) %>%
    select(record_id, date_value = date_stool_passing,redcap_repeat_instance,crf_completed_by_stool) %>%
    mutate(date_column = "date_stool_passing")
) %>%
  arrange(record_id, redcap_repeat_instance,crf_completed_by_stool,date_column)

write

save(early_dates_combined,file = paste0(file_path, "early_dates_combined.RData"))

```
### Future dates 
```{r}
today <- Sys.Date()  # today's date

future_dates_combined <- bind_rows(
  community_stool_df %>%
    filter(!is.na(date_stool) & as.Date(date_stool) > today) %>%
    select(record_id, date_value = date_stool,redcap_repeat_instance,crf_completed_by_stool) %>%
    mutate(date_column = "date_stool"),

  community_stool_df %>%
    filter(!is.na(date_consent_stool) & as.Date(date_consent_stool) > today) %>%
    select(record_id, date_value = date_consent_stool,redcap_repeat_instance,crf_completed_by_stool) %>%
    mutate(date_column = "date_consent_stool"),

  community_stool_df %>%
    filter(!is.na(date_collect_stool) & as.Date(date_collect_stool) > today) %>%
    select(record_id, date_value = date_collect_stool,redcap_repeat_instance,crf_completed_by_stool) %>%
    mutate(date_column = "date_collect_stool"),

  community_stool_df %>%
    filter(!is.na(date_stool_passing) & as.Date(date_stool_passing) > today) %>%
    select(record_id, date_value = date_stool_passing,redcap_repeat_instance,crf_completed_by_stool) %>%
    mutate(date_column = "date_stool_passing")
) %>%
  arrange(record_id, date_column,redcap_repeat_instance,crf_completed_by_stool)

save(future_dates_combined,file = paste0(file_path, "future_dates_combined.RData"))
```
### Stool sample collected but missing some stool dates 
```{r}
sample_dates_missing <- community_stool_df %>%
  filter(sample_collected_stool == "Yes") %>%
  filter(
    is.na(date_stool) |
    is.na(date_consent_stool) |
    is.na(date_collect_stool) |
    is.na(date_stool_passing)
  ) %>%
  select(record_id,sample_collected_stool,date_stool, date_consent_stool,date_collect_stool, date_stool_passing,redcap_repeat_instance,
         crf_completed_by_stool)

save(sample_dates_missing,file = paste0(file_path, "sample_dates_missing.RData"))
```
### Stool sample not collected but missing reasons
```{r}
missing_reason_stool_not_collected <- community_stool_df %>%
  filter(
    sample_collected_stool == "No" &
    (is.na(rson_no_sample_stool) | rson_no_sample_stool == "")
  ) %>%
  select(record_id, sample_collected_stool, rson_no_sample_stool,redcap_repeat_instance,crf_completed_by_stool)

save(missing_reason_stool_not_collected,file = paste0(file_path, "missing_reason_stool_not_collected.RData"))
```

### Stool kidms Checks
```{r}
### Checking STOOL 
# 1. Total stool records
total_redcap_stool <- nrow(community_stool_df)
total_kidms_stool <- nrow(community_stool_kidms_df) 


dup_stool_redcap <- community_stool_df %>%
  count(number_stool) %>%
  filter(n > 1) %>%
  nrow()

# 3. Matching study_ids
matching_stool <- inner_join(community_stool_df, community_stool_kidms_df, by = c("number_stool" = "serial_study_id"))

matching_stool <- matching_stool %>% 
  select(number_stool,everything())

matched_count <- nrow(matching_stool)

# 4. Non-matching study_ids
stool_redcap_only <- anti_join(community_stool_df, community_stool_kidms_df, by = c("number_stool" = "serial_study_id")) %>% nrow()

stool_kidms_only <- anti_join(community_stool_kidms_df, community_stool_df, by = c("serial_study_id" = "number_stool")) %>% nrow()

# 5. Summary table
summary_table_stool <- tibble(
  Source = c("community_stool_df", "community_stool_kidms_df"),
  Total_stool = c(total_redcap_stool, total_kidms_stool),
  Matching_study_ids = c(matched_count, matched_count),
  Unmatched_study_ids = c(stool_redcap_only, stool_kidms_only)
)

#stool in REDCAP but NOT in kidms
stool_redcap_only <- anti_join(community_stool_df, community_stool_kidms_df, by = c("number_stool" = "serial_study_id")) %>% 
  select(record_id,number_stool,everything())

## 1. stool Record IDs in redcap but not in kidms:
stool_in_redcap_not_in_kidms <- anti_join(community_stool_df, community_stool_kidms_df, by = c("number_stool" = "serial_study_id"))

stool_in_redcap_not_in_kidms <- stool_in_redcap_not_in_kidms %>% 
  dplyr::filter(!is.na(date_collect_stool)) %>% 
  select(record_id,number_stool,everything())

write_csv(stool_in_redcap_not_in_kidms,file = here("queries","stool_in_redcap_not_in_kidms.csv"))
```

## stool Record IDs in kidms but not in dredcap:
```{r}
stool_wrong_serial_study_id <- anti_join(community_stool_kidms_df, community_stool_df, by = c("serial_study_id" = "number_stool")) 

write_csv(stool_wrong_serial_study_id,file = here("queries","stool_in_kidms_not_in_redcap.csv"))

save(stool_wrong_serial_study_id,file = paste0(file_path, "stool_wrong_serial_study_id.RData"))
```


```{r}
# stool_wrong_serial_study_id <- community_stool_kidms_df %>%
#   dplyr::filter(serial_study_id %in% c("1-61B-14-3W-F","
# 1-120-409-W-F1","1-120-409-W-F3","1-120-409-W-F4","1-12A-153-W-F","1-12A-153-W-F2","1-12A-153-W-F3","1-12A-153-W-F5","1-12A-153-W-F6","1-12A-44-2-F","1-12G-553-2W-F","1-12G-553-2W-F","112A110W-F2","112A110W-F3","112H185W-F1",
# "112H185W-F2","112D282W-F1","
# 112D282W-F2","112N175W-F2","112A287-F1337","
# 112A287-F1","112A287-F2","112A287-F3","1-116G-478-W-F","1-113-428-6-W-F","112A110W-F3","1-116D-463-2",
# "1-116G-478-14-W","1-116G-478-W",
# "1-117L-115-1","1-118B-","1-120-409-W-F1","1-12A-153-W-F1",
# "1-161l-342-10","1-162D-59-7",
# "112035W-F1","112035W-F5","112035W-F6",
# "112A287-F1","112A287-F4","112D282W-F2","12035W-F2","12035W-F4","120617W-F1","120617W-F2","120617W-F3",
# "120617W-F4","1A005-W-F","1A006-W-F")) %>%
#   select(serial_study_id,pk_specimen_id,date_collect,type_name)
# 
# kbl(stool_wrong_serial_study_id,caption = "Wrong format of serial study id",booktabs = TRUE,
#     linesep = "\\addlinespace",align = "lcccc") %>%
#   kable_styling(latex_options = c("striped", "scale_down", "hold_position"),font_size = 8,
#     full_width = TRUE, position = "left") %>% row_spec(0, bold = TRUE)
# 
# write_csv(stool_wrong_serial_study_id, file = here("queries","stool_wrong_serial_study_id.csv"))
# 
# save(stool_wrong_serial_study_id,file = paste0(file_path, "stool_wrong_serial_study_id.RData"))
```

## Missing time_point
```{r}

miss_stool_time_point<-community_stool_kidms_df %>% 
  dplyr::filter(!time_point == "Wet") %>% 
  select(serial_study_id,pk_specimen_id,date_collect,time_point)

kbl(miss_stool_time_point,caption = "Missing stool time point ",booktabs = TRUE,
    linesep = "\\addlinespace",align = "lcccc") %>%
  kable_styling(latex_options = c("striped", "scale_down", "hold_position"),font_size = 8,
    full_width = TRUE, position = "left") %>% row_spec(0, bold = TRUE) 

write_csv(miss_stool_time_point, file = here("queries","miss_stool_time_point.csv"))

save(miss_stool_time_point,file = paste0(file_path, "miss_stool_time_point.RData"))
```

### Admission samples - Rectal swabs and stool

```{r}
admission_samples_summary <- hospital_df %>%
  filter(!is.na(cohort)) %>%
  group_by(cohort) %>%
  summarise(
    Rectal_swabs_collected = sum(!is.na(date_rectswab_adm)),
    Stool_samples_collected = sum(!is.na(date_stoolsample_adm)),
    .groups = "drop") %>%
  mutate(
    Target = case_when(
    cohort == "Adult" ~ 100,
    cohort == "Paeds" ~ 100),
    Total_samples = Rectal_swabs_collected + Stool_samples_collected) %>%
  select( cohort,Target,Rectal_swabs_collected,Stool_samples_collected,Total_samples) %>%
  adorn_totals(where = "row")


kbl(admission_samples_summary,caption = "Admission samples summary",booktabs = TRUE,
    linesep = "\\addlinespace",align = "lcccc") %>%
  kable_styling(latex_options = c("striped", "scale_down", "hold_position"),font_size = 8,
    full_width = TRUE, position = "left") %>% row_spec(0, bold = TRUE)

save(admission_samples_summary,file = paste0(file_path, "admission_samples_summary.RData") )

```
### Rectal swab checks 
```{r}
R1_redcap <- hospital_df %>% 
  dplyr::filter(!is.na(sample_num_rectswab1_adm)) %>% 
  select(record_id, sample_num_rectswab1_adm,rectswab_adm,cohort)

adm_stool <- hospital_df %>% 
  dplyr::filter(!is.na(sample_num_stool_adm)) %>% 
  select(record_id,sample_num_stool_adm,stoolsample_adm)

R1_swabs_kidms <- rectal_swabs_kidms_df %>%
  filter(str_detect(serial_study_id, "-R1$")) 
  
all_R1_swabs <- left_join(R1_redcap, R1_swabs_kidms, by = c("sample_num_rectswab1_adm" = "serial_study_id"))

# 1. Total R1 records
total_redcap <- nrow(R1_redcap)
total_kidms <- nrow(R1_swabs_kidms) 

# 2. Duplicated study_ids
dup_redcap <- R1_redcap %>%
  count(sample_num_rectswab1_adm) %>%
  filter(n > 1) %>%
  nrow()

dup_kidms <- R1_swabs_kidms %>%
  count(serial_study_id) %>%
  filter(n > 1) %>%
  nrow()
# 3. Matching study_ids
matching_R1 <- inner_join(R1_redcap, R1_swabs_kidms, by = c("sample_num_rectswab1_adm" = "serial_study_id"))
matched_count <- nrow(matching_R1)

# 4. Non-matching study_ids
redcap_only <- anti_join(R1_redcap, R1_swabs_kidms, by = c("sample_num_rectswab1_adm" = "serial_study_id")) %>% nrow()

kidms_only <- anti_join(R1_swabs_kidms, R1_redcap, by = c("serial_study_id" = "sample_num_rectswab1_adm")) %>% nrow()

# 5. Summary table
summary_table <- tibble(
  Source = c("R1_redcap", "R1_swabs_kidms"),
  Total_R1s = c(total_redcap, total_kidms),
  Duplicated_study_ids = c(dup_redcap, dup_kidms),
  Matching_study_ids = c(matched_count, matched_count),
  Unmatched_study_ids = c(redcap_only, kidms_only)
)

#R1 swabs in REDCAP but NOT in kidms
R1_redcap_only <- anti_join(R1_redcap, R1_swabs_kidms, by = c("sample_num_rectswab1_adm" = "serial_study_id")) %>% 
  select(record_id,sample_num_rectswab1_adm,everything())

write_csv(R1_redcap_only,file = here("queries","R1_redcap_only.csv"))

save(R1_redcap_only,file = paste0(file_path, "R1_redcap_only.RData")) 
```

```{r}
rectswab1_miss <- rectal_swabs_kidms_df %>% 
  dplyr::filter(serial_study_id %in% c("1P002-W-R2","1P002-W-R2","1P003-W-R2","1P003-W-R2")) %>% 
  select(serial_study_id,type_name,date_collect,pk_specimen_id)
 
kbl(rectswab1_miss,caption = "Record ids missing rectal swab 1",booktabs = TRUE,
    linesep = "\\addlinespace",align = "lcccc") %>%
  kable_styling(latex_options = c("striped", "scale_down", "hold_position"),font_size = 8,
    full_width = TRUE, position = "left") %>% row_spec(0, bold = TRUE) 

save(rectswab1_miss,file = paste0(file_path, "rectswab1_miss.RData")) 

### wrong serial number format 
wrong_serial_study_id_format <- rectal_swabs_kidms_df %>% 
  dplyr::filter(serial_study_id %in% c("1P-023-W-R2","1P026","IP027-W-R1","IP027-W-R2
","1PO42-W-R2","1P060")) %>% 
  select(serial_study_id,type_name,date_collect,pk_specimen_id)
  
kbl(wrong_serial_study_id_format,caption = "Wrong format of serial study id",booktabs = TRUE,
    linesep = "\\addlinespace",align = "lcccc") %>%
  kable_styling(latex_options = c("striped", "scale_down", "hold_position"),font_size = 8,
    full_width = TRUE, position = "left") %>% row_spec(0, bold = TRUE) 

write_csv(wrong_serial_study_id_format, file = here("queries","wrong_serial_study_id_format.csv"))

save(wrong_serial_study_id_format,file = paste0(file_path, "wrong_serial_study_id_format.RData")) 

### Rectal swab specimen source 
rectswab_specimen_source <- rectal_swabs_kidms_df %>% 
  dplyr::filter(specimen_source %in% c("clinic","field")) %>% 
  select(serial_study_id,pk_specimen_id,specimen_source,type_name,date_collect)

kbl(rectswab_specimen_source,caption = "Wrong rectal_swab specimen source",booktabs = TRUE,
    linesep = "\\addlinespace",align = "lcccc") %>%
  kable_styling(latex_options = c("striped", "scale_down", "hold_position"),font_size = 8,
    full_width = TRUE, position = "left") %>% row_spec(0, bold = TRUE) 

write_csv(rectswab_specimen_source, file = here("queries","rectswab_specimen_source.csv"))

save(rectswab_specimen_source,file = paste0(file_path, "rectswab_specimen_source.RData")) 

## Rectal swabs missing date recieved 

rectswab_date_recieved_miss <- rectal_swabs_kidms_df %>%
  dplyr::filter(is.na(date_received)) %>% 
  select(serial_study_id, type_name,pk_specimen_id,date_collect, date_received)
  
kbl(rectswab_date_recieved_miss,caption = "Rectswab missing date_recieved in the lab",booktabs = TRUE,
    linesep = "\\addlinespace",align = "lcccc") %>%
  kable_styling(latex_options = c("striped", "scale_down", "hold_position"),font_size = 8,
    full_width = TRUE, position = "left") %>% row_spec(0, bold = TRUE) 

write_csv(rectswab_date_recieved_miss, file = here("queries","rectswab_date_recieved_miss.csv"))

save(rectswab_date_recieved_miss,file = paste0(file_path, "rectswab_date_recieved_miss.RData"))

## Rectal swabs missing time recieved 

rectswab_time_recieved_miss <- rectal_swabs_kidms_df %>%
  dplyr::filter(is.na(time_received)) %>% 
  select(serial_study_id, type_name,pk_specimen_id,date_collect, time_received)
  
kbl(rectswab_time_recieved_miss,caption = "Rectswab missing time_recieved in the lab",booktabs = TRUE,
    linesep = "\\addlinespace",align = "lcccc") %>%
  kable_styling(latex_options = c("striped", "scale_down", "hold_position"),font_size = 8,
    full_width = TRUE, position = "left") %>% row_spec(0, bold = TRUE) 

write_csv(rectswab_time_recieved_miss, file = here("queries","rectswab_time_recieved_miss.csv"))

save(rectswab_time_recieved_miss,file = paste0(file_path, "rectswab_time_recieved_miss.RData"))

```
### Environmental samples  

```{r}
# environmental_sample_summary <- community_wet_df %>%
#   summarise(
#     Total_homesteads = sum(!is.na(date_survey_hh), na.rm = TRUE), 
#     Water_samples = sum(!is.na(date_collect_ws), na.rm = TRUE),
#     Soil_samples = sum(!is.na(date_collect_ss), na.rm = TRUE),
#     Latrine_samples = sum(!is.na(date_collect_ls), na.rm = TRUE),
#     Outdoor_samples = sum(!is.na(date_collect_ls_2), na.rm = TRUE),
#     Chicken_faeces_samples = sum(!is.na(date_collect_cf), na.rm = TRUE),
#     
#     Total_samples = Water_samples + Soil_samples + Latrine_samples + Outdoor_samples + Chicken_faeces_samples
#   )
# save(environmental_sample_summary,file = paste0(file_path, "environmental_sample_summary.RData") )


```
### Environmental samples 2

```{r}
environmental_sample_summary_2 <- community_wet_df %>%
  summarise(
    Total_homesteads = sum(!is.na(date_survey_hh), na.rm = TRUE), 
    
    Water_samples = paste0(
      sum(!is.na(date_collect_ws), na.rm = TRUE), " (",
      round(sum(!is.na(date_collect_ws), na.rm = TRUE) / Total_homesteads * 100, 1), "%)"
    ),
    
    Soil_samples = paste0(
      sum(!is.na(date_collect_ss), na.rm = TRUE), " (",
      round(sum(!is.na(date_collect_ss), na.rm = TRUE) / Total_homesteads * 100, 1), "%)"
    ),
    
    Latrine_samples = paste0(
      sum(!is.na(date_collect_ls), na.rm = TRUE), " (",
      round(sum(!is.na(date_collect_ls), na.rm = TRUE) / Total_homesteads * 100, 1), "%)"
    ),
    
    Outdoor_samples = paste0(
      sum(!is.na(date_collect_ls_2), na.rm = TRUE), " (",
      round(sum(!is.na(date_collect_ls_2), na.rm = TRUE) / Total_homesteads * 100, 1), "%)"
    ),
    
    Chicken_faeces_samples = paste0(
      sum(!is.na(date_collect_cf), na.rm = TRUE), " (",
      round(sum(!is.na(date_collect_cf), na.rm = TRUE) / Total_homesteads * 100, 1), "%)"
    )
  ) %>%
  pivot_longer(
    cols = -Total_homesteads,
    names_to = "Sample_Type",
    values_to = "Sample_Count_%"
  )
save(environmental_sample_summary_2,file = paste0(file_path, "environmental_sample_summary_2.RData"))
```
### Reasons Latrine sample not collected 
```{r}
rsns_latrine_not_collected <- community_wet_df %>%
  filter(latrine_sample_collected_ls == "No") %>%
  group_by(rson_no_sample_ls) %>%
  summarise(Total = n(), .groups = "drop") %>%
  mutate(
    Percent = round((Total / sum(Total)) * 100, 1)
  ) %>%
  mutate(
    Percent = paste0(Percent, "%")
  ) %>% 
  adorn_totals(where = "row")

save(rsns_latrine_not_collected,file = paste0(file_path, "rsns_latrine_not_collected.RData"))
```
### Latrine and outdoor samples summary
```{r}
# Step 1: Summarise main sample group counts
combined_summary <- community_wet_df %>%
  mutate(
    Sample_group = case_when(
      latrine_sample_collected_ls == "Yes" ~ "Latrine Sample Collected",
      latrine_sample_collected_ls == "No" & outdoor_sample_collected_ls == "Yes" ~ "Latrine Sample Not Collected + Outdoor Sample Collected",
      latrine_sample_collected_ls == "No" & outdoor_sample_collected_ls == "No" ~ "Latrine Sample Not Collected + Outdoor Sample Not Collected",
      latrine_sample_collected_ls == "No" & is.na(outdoor_sample_collected_ls) ~ "Latrine Sample Not Collected + Outdoor Sample N/A",
      TRUE ~ "Other / Unknown"
    )
  ) %>%
  count(Sample_group, name = "Total")

# Step 2: Reason summaries for specific groups

# 2a: Latrine sample not collected + Outdoor sample collected
latrine_reasons_collected <- community_wet_df %>%
  filter(latrine_sample_collected_ls == "No", outdoor_sample_collected_ls == "Yes") %>%
  count(rson_no_sample_ls, name = "n") %>%
  mutate(reason_text = paste0(rson_no_sample_ls, " (", n, ")")) %>%
  summarise(text = paste(reason_text, collapse = "; ")) %>%
  pull(text)

# 2b: Latrine sample not collected + Outdoor sample not collected
latrine_reasons_not_collected_outdoor_not_collected <- community_wet_df %>%
  filter(latrine_sample_collected_ls == "No", outdoor_sample_collected_ls == "No", !is.na(rson_no_sample_ls)) %>%
  count(rson_no_sample_ls, name = "n") %>%
  mutate(reason_text = paste0(rson_no_sample_ls, " (", n, ")")) %>%
  summarise(text = paste(reason_text, collapse = "; ")) %>%
  pull(text)

# 2c: Latrine sample not collected + Outdoorsample is blank, excluding "No facility"
latrine_reasons_blank <- community_wet_df %>%
  filter(latrine_sample_collected_ls == "No", is.na(outdoor_sample_collected_ls),
         !is.na(rson_no_sample_ls), rson_no_sample_ls != "No facility") %>%
  count(rson_no_sample_ls, name = "n") %>%
  mutate(reason_text = paste0(rson_no_sample_ls, " (", n, ")")) %>%
  summarise(text = paste(reason_text, collapse = "; ")) %>%
  pull(text)

# 2d: Outdoor sample not collected reasons
outdoor_reasons <- community_wet_df %>%
  filter(outdoor_sample_collected_ls == "No", !is.na(rson_no_sample_od_ls)) %>%
  count(rson_no_sample_od_ls, name = "n") %>%
  mutate(reason_text = paste0(rson_no_sample_od_ls, " (", n, ")")) %>%
  summarise(text = paste(reason_text, collapse = "; ")) %>%
  pull(text)

# Step 3: Add all reasons into final summary
latrine_outdoor_summary <- combined_summary %>%
  mutate(
    Latrine_Reasons = case_when(
      Sample_group == "Latrine Sample Not Collected + Outdoor Sample Collected" ~ latrine_reasons_collected,
      Sample_group == "Latrine Sample Not Collected + Outdoor Sample N/A" ~ latrine_reasons_blank,
      Sample_group == "Latrine Sample Not Collected + Outdoor Sample Not Collected" ~ latrine_reasons_not_collected_outdoor_not_collected,
      TRUE ~ ""
    ),
    Outdoor_Reasons = case_when(
      Sample_group == "Latrine Sample Not Collected + Outdoor Sample Not Collected" ~ outdoor_reasons,
      TRUE ~ ""
    )
  ) %>%
  adorn_totals(where = "row")

write_csv(latrine_outdoor_summary, file = here("data","latrine_outdoor_summary.csv"))

save(latrine_outdoor_summary,file = paste0(file_path, "latrine_outdoor_summary.RData"))

```

```{r}
rsns_latrine_not_collected <- community_wet_df %>%
  filter(latrine_sample_collected_ls == "No") %>% 
  select(record_id,latrine_sample_collected_ls,rson_no_sample_ls,outdoor_sample_collected_ls,rson_no_sample_od_ls,latrine_sample_by_ls,outdoor_sample_by_ls)

write_csv(rsns_latrine_not_collected, here("data","rsns_latrine_not_collected.csv"))
```

```{r}
### Stool records 
records_stool_not_collected <- community_stool_df %>%
  filter(
    sample_collected_stool == "No") %>%
  select(
    record_id,redcap_repeat_instance,
    sample_collected_stool,
    rson_no_sample_stool,
    everything(),-c(redcap_event_name,redcap_repeat_instrument))

write_csv(records_stool_not_collected, file = here("data","records_stool_not_collected.csv"))
```
### Latrine records 

```{r}
latrine_not_collected_outdoor_missing <- community_wet_df %>%
  filter(
    latrine_sample_collected_ls == "No",
    is.na(outdoor_sample_collected_ls)
  ) %>%
  select(
    record_id,
    latrine_sample_collected_ls,
    rson_no_sample_ls,  # Reason latrine not collected
    date_collect_ls,
    time_collect_ls,
    outdoor_sample_collected_ls,
    rson_no_sample_od_ls,  # Reason outdoor not collected (may also be NA)
    
  )
latrine_not_collected_outdoor_missing$hm_number_hh <- sapply(
  strsplit(latrine_not_collected_outdoor_missing$record_id, "-"),
  function(x) x[3]
)

latrine_not_collected_outdoor_missing <- latrine_not_collected_outdoor_missing %>%
  select(record_id, hm_number_hh, everything())
write_csv(latrine_not_collected_outdoor_missing,file =here("data","latrine_not_collected_outdoor_missing.csv"))
```

```{r}
latrine_outdoor_not_collected <- community_wet_df %>%
  filter(
    latrine_sample_collected_ls == "No",
    outdoor_sample_collected_ls == "No"
  ) %>%
  select(
    record_id,
    latrine_sample_collected_ls,
    rson_no_sample_ls,
    date_collect_ls,
    time_collect_ls,
    outdoor_sample_collected_ls,
    rson_no_sample_od_ls,
    outdoor_sample_number_ls
  )

write_csv(latrine_outdoor_not_collected,file =here("data","latrine_outdoor_not_collected.csv"))
```

```{r}

stool = community_stool_df %>% 
  select(record_id,date_stool,sample_collected_stool,rson_no_sample_stool,date_collect_stool,number_stool)

nolatrine_withstool <- latrine_outdoor_not_collected %>% 
  left_join(stool, by= "record_id") 

nolatrine_withstool <- nolatrine_withstool %>% 
  select(record_id, number_stool, everything()) %>% 
  dplyr::filter(is.na(number_stool))


```
### Latrine and outdoor not collected 
```{r}
miss_latrine_outdoor <- community_wet_df %>% 
  dplyr::filter(latrine_sample_collected_ls == "No") %>% 
  dplyr::filter(outdoor_sample_collected_ls == "No") %>% 
select(record_id,latrine_sample_collected_ls,rson_no_sample_ls,date_collect_ls,latrine_sample_number_ls,outdoor_sample_collected_ls,rson_no_sample_od_ls,date_collect_ls_2,outdoor_sample_number_ls)

stool = community_stool_df %>% 
  select(record_id,date_stool,sample_collected_stool,rson_no_sample_stool,date_collect_stool,number_stool)

miss_latrine_outdoor <-  miss_latrine_outdoor %>% 
  left_join(stool, by = "record_id")

miss_latrine_outdoor <- miss_latrine_outdoor %>% 
   select(record_id, number_stool, everything()) %>% 
   dplyr::filter(is.na(number_stool))

miss_latrine_outdoor <- miss_latrine_outdoor %>% 
  filter(is.na(number_stool)) %>%
  distinct(record_id, .keep_all = TRUE)
  
write_csv(miss_latrine_outdoor,file = here("data_cleaning","miss_stool_latrine_outdoor1.csv")) 
```
### Latrine not collected , outdoor missing 
```{r}
latrine_not_collected_outdoor_missing <- community_wet_df %>%
  filter(
    latrine_sample_collected_ls == "No",
    is.na(outdoor_sample_collected_ls)
  ) %>%
  select(
    record_id,
    latrine_sample_collected_ls,
    rson_no_sample_ls,  # Reason latrine not collected
    date_collect_ls,
    time_collect_ls,
    outdoor_sample_collected_ls,
    rson_no_sample_od_ls,  # Reason outdoor not collected (may also be NA)
    
  )

latrine_not_collected_outdoor_missing <- latrine_not_collected_outdoor_missing %>% 
  left_join(stool, by = "record_id")

latrine_not_collected_outdoor_missing <- latrine_not_collected_outdoor_missing %>% 
   select(record_id, number_stool, everything()) %>% 
   dplyr::filter(is.na(number_stool)) 

latrine_not_collected_outdoor_missing <- latrine_not_collected_outdoor_missing %>%
  filter(is.na(number_stool)) %>%
  distinct(record_id, .keep_all = TRUE)

write_csv(latrine_not_collected_outdoor_missing,file = here("data_cleaning","miss_stool_latrine_outdoor2.csv"))
```

### PPS summaries

```{r}
## Join pps ward and pps patient 
pps_ward_df <- pps_ward_df %>% 
  select(-redcap_event_name)
pps_all<- full_join(pps_ward_df, pps_patient_df, by = "record_id")


pps_patient_summary <- pps_all %>%
  mutate(ward_round_pps = paste0("Round", ward_round_pps)) %>%
  group_by(ward_name_pps, ward_round_pps) %>%
  summarise(
    total_patients = n_distinct(part_id_pps),
    .groups = "drop"
  ) %>%
  pivot_wider(
    names_from = ward_round_pps,
    values_from = total_patients,
    values_fill = 0
  ) %>%
  rowwise() %>%
  mutate(Total_Patients = sum(c_across(starts_with("Round")))) %>%
  ungroup() %>%
  arrange(desc(ward_name_pps)) %>%
  adorn_totals(where = "row") %>% 
  rename(Ward_Name = ward_name_pps) 
 

kbl(pps_patient_summary,caption = "PPS Summary",booktabs = TRUE,
    linesep = "\\addlinespace",align = "lcccc") %>%
  kable_styling(latex_options = c("striped", "scale_down", "hold_position"),font_size = 8,
    full_width = TRUE, position = "left") %>% row_spec(0, bold = TRUE) 

save(pps_patient_summary,file = paste0(file_path, "pps_patient_summary.RData"))
```
### Exit summaries 
```{r}
exit_balance <- exit_df %>% 
  group_by(cluster_exit) %>% 
summarise(total = n())

write_csv(exit_balance,file = here("field_work_register","exit_balance.csv"))

missing_cluster <- exit_df %>% 
  dplyr::filter(is.na(cluster_exit)) %>% 
  select(record_id,cluster_exit,date_interview_exit,crf_completed_by_exit)

write_csv(missing_cluster,file = here("field_work_register","missing_cluster.csv"))
```
 ### Checking STOOL 
```{r}
# 1. Total stool records
total_redcap_stool <- nrow(community_stool_df)
total_kidms_stool <- nrow(community_stool_kidms_df) 


dup_stool_redcap <- community_stool_df %>%
  count(number_stool) %>%
  filter(n > 1) %>%
  nrow()

# 3. Matching study_ids
matching_stool <- inner_join(community_stool_df, community_stool_kidms_df, by = c("number_stool" = "serial_study_id"))

matching_stool <- matching_stool %>% 
  select(number_stool,everything())

matched_count <- nrow(matching_stool)

# 4. Non-matching study_ids
stool_redcap_only <- anti_join(community_stool_df, community_stool_kidms_df, by = c("number_stool" = "serial_study_id")) %>% nrow()

stool_kidms_only <- anti_join(community_stool_kidms_df, community_stool_df, by = c("serial_study_id" = "number_stool")) %>% nrow()

# 5. Summary table
summary_table_stool <- tibble(
  Source = c("community_stool_df", "community_stool_kidms_df"),
  Total_stool = c(total_redcap_stool, total_kidms_stool),
  Matching_study_ids = c(matched_count, matched_count),
  Unmatched_study_ids = c(stool_redcap_only, stool_kidms_only)
)

#stool in REDCAP but NOT in kidms
stool_redcap_only <- anti_join(community_stool_df, community_stool_kidms_df, by = c("number_stool" = "serial_study_id")) %>% 
  select(record_id,number_stool,everything())

## 1. stool Record IDs in redcap but not in kidms:
stool_in_redcap_not_in_kidms <- anti_join(community_stool_df, community_stool_kidms_df, by = c("number_stool" = "serial_study_id"))

stool_in_redcap_not_in_kidms <- stool_in_redcap_not_in_kidms %>% 
  dplyr::filter(!is.na(date_collect_stool)) %>% 
  select(record_id,number_stool,everything())

write_csv(stool_in_redcap_not_in_kidms,file = here("queries","stool_in_redcap_not_in_kidms.csv"))
```

## stool Record IDs in kidms but not in dredcap:
```{r}
stool_in_kidms_not_in_redcap <- anti_join(community_stool_kidms_df, community_stool_df, by = c("serial_study_id" = "number_stool")) %>% 
  distinct(serial_study_id,.keep_all = TRUE)

write_csv(stool_in_kidms_not_in_redcap,file = here("queries","stool_in_kidms_not_in_redcap.csv"))
```

